---
title: "csd_eda"
author: "Jerry_Chen"
date: "November 20, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(dplyr)
library(car)
```

## Preprocessing & Cleaning

```{r}
csd <- read_csv("catalog sales data.csv", col_types = cols(datead6 = col_date(format = "%m/%d/%Y"), datelp6 = col_date(format = "%m/%d/%Y")))

#summary(csd)
head(csd)
```

```{r}
csd <- subset(csd, select = -c(ordhist) )
```


```{r}
csd[is.na(csd$lpuryear) == FALSE & 
      substr(csd$datelp6,4,4) < csd$lpuryear,] <- csd %>% 
  filter(is.na(lpuryear) == FALSE & substr(datelp6,4,4) < lpuryear) %>%
  mutate(datelp6 = as.Date(paste0(substr(datelp6,1,3), lpuryear, substr(datelp6,5,10))))

csd
```

```{r}
train <- csd[csd$train == 1,]
test <- csd[csd$train == 0,]
```

```{r}
train$targ_prob <- ifelse(train$targdol > 0, 1, 0)
table(train$targ_prob)
```

## EDA

select numerical features and check correlation
```{r}
train_num <- train %>% select(-datead6, -datelp6, -lpuryear, -train)
cor(train_num)
```

###A few comments:
  1. LTD (slshist) may be correlated with other sales, and also with fall order and spring order. (It is better to dropp slshist.)
  2. sales are not highly correlated with each other. The same goes with orders.
  3. the sales and order are highly correlated during the same period, e.g., slstyr with ordtyr, sls2ago with ord2ago.

Regression of historical sales (LTD) onto other sales (this year, last year, two years, three years)
```{r}
sls_lm <- lm(slshist ~ slstyr+slslyr+sls2ago+sls3ago, data=train)
summary(sls_lm)
```

sales are not highly correlated with each other, except with the historical sales.
calculate the VIFs of the simple linear regression.
```{r}
vif(sls_lm)
```

Repeat this on order variables:
```{r}
ord_lm <- lm(falord ~ ordtyr+ordlyr+ord2ago+ord3ago, data=train)
summary(ord_lm)
```

```{r}
vif(ord_lm)
```


Try to use logistic regression (using order variables) to predict the probability of purchase.
```{r}
targ_prob_ord_log <- glm(targ_prob~ordtyr+ordlyr+ord2ago+ord3ago, data=train, family=binomial)
summary(targ_prob_ord_log)
```

```{r}
conf_ord <- table(train$targ_prob, targ_prob_ord_log$fitted.values > 0.5)
conf_ord
```


Try to use logistic regression (using sales variables) to predict the probability of purchase.
```{r}
targ_prob_sls_log <- glm(targ_prob~slstyr+slslyr+sls2ago+sls3ago, data=train, family=binomial)
summary(targ_prob_sls_log)
```
```{r}
conf_sls <- table(train$targ_prob, targ_prob_sls_log$fitted.values > 0.5)
conf_sls
```

Use F-1 score because the data set is not balanced.
