---
title: "Predictive1_Project_Part1"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
```{r}
library(tidyverse)
library(glmnet)
library(ordinal)
library(mlogit)
library(car)
```

## Data Cleansing

We took 2 main steps to clean the data:
1) Change ordhist value to the sum of falord and sprord
2) Update the datelp6 year to match lpuryear where needed
3) Drop lpuryear

```{r}
csd <- read_csv("~/z/homework/pd_project/Predictive1_Proj_Champs/catalog sales data.csv", 
                col_types = cols(datead6 = col_date(format = "%m/%d/%Y"), 
                                 datelp6 = col_date(format = "%m/%d/%Y")))

#Drop ordhist
#csd <- subset(csd, select = -c(ordhist))

#create new ordhist by adding falord and sprord
csd$ordhist <- csd$falord + csd$sprord

#Update daltelp6
csd[is.na(csd$lpuryear) == FALSE & substr(csd$datelp6,4,4) < csd$lpuryear & as.numeric(paste0(substr(csd$datelp6,3,3),csd$lpuryear)) < 13,] <- csd %>% 
  filter(is.na(lpuryear) == FALSE & substr(datelp6,4,4) < lpuryear & as.numeric(paste0(substr(datelp6,3,3),lpuryear)) < 13) %>%
  mutate(datelp6 = as.Date(paste0(substr(datelp6,1,3), lpuryear,"-06-30")))

#Drop lpuryear
csd <- subset(csd, select = -c(lpuryear))
```

```{r}
mis_sls <- csd %>%
  mutate(oldSales = slshist - (slstyr + slslyr + sls2ago + sls3ago)) %>%
  filter(oldSales != 0) %>%
  filter(as.integer(substr(datelp6,1,4)) - 3 < as.integer(substr(datead6,1,4))) %>%
  select(-targdol, -train)
sum(mis_sls$oldSales)
```

```{r}
mis_ord <- csd %>%
  mutate(oldSales = sprord + falord - (ordtyr + ordlyr + ord2ago + ord3ago)) %>%
  filter(oldSales != 0) %>%
  filter(as.integer(substr(datelp6,1,4)) - 3 < as.integer(substr(datead6,1,4))) %>%
  select(-targdol, -train)
sum(mis_ord$oldSales)
```

Adding columns for sales and orders more than 3 years ago
```{r}
new_csd <- csd %>%
  mutate(oldOrds = sprord + falord - (ordtyr + ordlyr + ord2ago + ord3ago)) %>%
  mutate(oldSales = slshist - (slstyr + slslyr + sls2ago + sls3ago)) %>%
  #filter(!((as.integer(substr(datelp6,1,4)) - 3 <= as.integer(substr(datead6,1,4))) & (oldOrds != 0) & (oldSales !=0))) %>%
  mutate(oldOrds = ifelse(oldOrds < 0, 0, oldOrds))
```

EDA histograms
```{r}
hist(csd$slstyr)
hist(csd$slslyr)
hist(csd$sls2ago)
hist(csd$sls3ago)
hist(csd$ordtyr)
hist(csd$ordlyr)
hist(csd$ord2ago)
hist(csd$ord3ago)
```

EDA graphs
```{r}
new_csd %>%
  filter(targdol != 0) %>%
  ggplot(aes(slstyr)) + 
  geom_histogram(binwidth = .25) + 
  labs(title = "sales ($) this year")

new_csd %>%
  filter(targdol != 0) %>%
  ggplot(aes(ordtyr)) + 
  geom_histogram(binwidth = .25) + 
  labs(title = "orders this year")

new_csd %>%
  filter(targdol != 0) %>%
  ggplot(aes(ordlyr)) + 
  geom_histogram(binwidth = .25) + 
  labs(title = "orders last year")

new_csd %>%
  filter(targdol != 0) %>%
  ggplot(aes(ord2ago)) + 
  geom_histogram(binwidth = .25) + 
  labs(title = "orders 2 years ago")

new_csd %>%
  filter(targdol != 0) %>%
  ggplot(aes(ord3ago)) + 
  geom_histogram(binwidth = .25) + 
  labs(title = "orders 3 years ago")

new_csd %>%
  filter(targdol != 0) %>%
  ggplot(aes(ordhist)) + 
  geom_histogram(binwidth = .25) + 
  labs(title = "LTD orders")

new_csd %>%
  filter(targdol != 0) %>%
  ggplot(aes(slshist)) + 
  geom_histogram(binwidth = .25) +
  labs(title = "LTD dollars")

new_csd %>%
  filter(targdol != 0) %>%
  ggplot(aes(oldSales)) + 
  geom_histogram(binwidth = .25) +
  labs(title = ">3yr old dollars")

new_csd %>%
  filter(targdol != 0) %>%
  ggplot(aes(oldOrds)) + 
  geom_histogram(binwidth = .25) +
  labs(title = ">3yr old ords")
```

Add recency as in days data
```{r}
new_csd <- new_csd %>%
  mutate(recency = (2017-as.integer(substr(datelp6,1,4)))*12*30 + (11 - as.integer(substr(datelp6,6,7)))*30 + (20-as.integer(substr(datelp6,9,10))))
```

Change recency from days based to year based
```{r}
new_csd$recency <- new_csd$recency/365
```

```{r}
hist(new_csd$recency)
hist(log(new_csd$recency))
```

Prep the data for binomial prediction, adding targdec column (0/1)
```{r}
new_csd <- new_csd %>%
  mutate(targdec = ifelse(targdol > 0, 1, 0)) %>%
  select(-datead6, datelp6)
```

##Treat consistency as 1/0 for each year if needed
new_csd <- new_csd %>%
  mutate(yr0 = ifelse (slstyr > 0, 1, 0)) %>%
  mutate(yr1 = ifelse (slslyr > 0, 1, 0)) %>%
  mutate(yr2 = ifelse (sls2ago > 0, 1, 0)) %>%
  mutate(yr3 = ifelse (sls3ago > 0, 1, 0)) %>%
  mutate(yr4 = ifelse (oldSales > 0, 1, 0)) %>%
  mutate(constcy = yr0 + yr1 + yr2 + yr3 + yr4) %>%
  select(-yr0, -yr1, -yr2, -yr3, -yr4)

Reading in the data and separating into training and test sets.
```{r}
#catalog <- read.csv("catalog sales data.csv", header = T)
newtrain <- new_csd[new_csd$train == 1,][,-which(names(new_csd) == 'train')]
newtest <- new_csd[new_csd$train == 0,][,-which(names(new_csd) == 'train')]
```

##Ordinal model may not be needed
newtrain$targdec.ordered = as.ordered(newtrain$targdec)
fit = clm(targdec.ordered~constcy+recency, data = newtrain)
summary(fit)


#prep data for logistic model and testing
```{r}
newtrain1 <- mlogit.data(data = newtrain, choice = "targdec", shape = "wide", varying = NULL)
newtest1 <- mlogit.data(data = newtest, choice = "targdec", shape = "wide", varying = NULL)
```

#Logistic Models
##Logistic Model without interactions
```{r}
fit1 <- mlogit(targdec~0|ordtyr+ordlyr+ord2ago+ord3ago+oldSales+oldOrds+recency, data=newtrain1)
summary(fit1)
```

```{r}
testpred1 <- predict(fit1, newtest1)
testpred1 <- as.integer(colnames(testpred1)[apply(testpred1,1,which.max)])
```

```{r}
cattable1 = table(newtest$targdec, testpred1)
cattable1
sum(diag(cattable1))/sum(cattable1)
```

F1 score
```{r}
P <- cattable1[2,2]/(sum(cattable1[2,]))
R <- cattable1[2,2]/(sum(cattable1[,2]))
2*P*R/(P+R)
```

##Logistic Model with Normal Data
```{r}
fit2 <- mlogit(targdec~0|ordtyr+ordlyr+ord2ago+ord3ago+oldSales+oldOrds+recency+ordtyr:ordlyr:ord2ago:ord3ago:oldOrds, data=newtrain1)
summary(fit2)
```

```{r}
newtest1 <- mlogit.data(data = newtest, choice = "targdec", shape = "wide", varying = NULL)
testpred <- predict(fit2, newtest1)
testpred <- as.integer(colnames(testpred)[apply(testpred,1,which.max)])
```

```{r}
cattable = table(newtest$targdec, testpred)
cattable
sum(diag(cattable))/sum(cattable)
```

#Multiple regression model
```{r}
mlmtrain <- newtrain[newtrain$targdec == 1,][,-which(names(newtrain) == 'targdec')]
```

filter only targdol>0 & add recent sales and recent orders columns for train and test
```{r}
mlmtrain <- mlmtrain %>%
  select(-datelp6,-falord,-sprord) %>%
  filter(targdol > 0) %>%
  mutate(slsrecent = slstyr+slslyr+sls2ago+sls3ago) %>%
  mutate(ordrecent = ordtyr+ordlyr+ord2ago+ord3ago)
```

```{r}
mlmtest <- newtest %>%
  filter(targdol > 0) %>%
  select(-datelp6,-falord,-sprord) %>%
  mutate(slsrecent = slstyr+slslyr+sls2ago+sls3ago) %>%
  mutate(ordrecent = ordtyr+ordlyr+ord2ago+ord3ago)
```

##full multi regression model
```{r}
mlmfit <- lm(targdol~., data = mlmtrain)
summary(mlmfit)
sum(((mlmtest$targdol) - (predict(mlmfit, newdata = mlmtest)))^2)
plot(mlmfit)
```

##full mlm with log transformation
```{r}
mlmfit.log <- lm(log(targdol)~., data = mlmtrain)
summary(mlmfit.log)
sum(((mlmtest$targdol) - (predict(mlmfit.log, newdata = mlmtest)))^2)
plot(mlmfit.log)
```

##full mlm with box cox transformation
```{r}
boxCox(mlmfit, family="yjPower", plotit = TRUE)
#tranform response
depvar.transformed <- yjPower(mlmtrain$targdol, -.2)
glm.mod <- glm(depvar.transformed~+slstyr+slslyr+sls2ago+sls3ago+slshist+ordtyr+ordlyr+ord2ago+ord3ago+ordhist+oldOrds+oldSales+recency+slsrecent+ordrecent, family=gaussian(link="log"), data = mlmtrain)
summary(glm.mod)
sum(((mlmtest$targdol) - (predict(glm.mod, newdata = mlmtest)))^2)
plot(glm.mod)
#transform predictors
#boxTidwell(depvar.transformed~slstyr:slslyr:sls2ago:sls3ago:slshist:oldSales+ordtyr:ordlyr:ord2ago:ord3ago:ordhist:oldOrds+recency+slsrecent:ordrecent, data=mlmtrain)
```

##steps with box cosx

##steps with no transformation
```{r}
step(mlmfit, direction = "backward")
```
##best fit using backward steps
```{r}
mlmfit.steps <- lm(targdol ~ slstyr + slslyr + sls2ago + slshist + ordtyr + ordlyr + 
    ord2ago + ord3ago, data=mlmtrain)
summary(mlmfit.steps)
sum(((mlmtest$targdol) - (predict(mlmfit.steps, newdata = mlmtest)))^2)
```

##steps using log transformation
```{r}
step(mlmfit.log, direction = "backward")
```

##best fit using steps with log transformation
```{r}
mlmfit.log.steps <- lm(log(targdol)~slstyr + slslyr + sls2ago + slshist + 
    ordtyr + ordlyr + ord2ago + ord3ago, data = mlmtrain)
summary(mlmfit.log.steps)
sum(((mlmtest$targdol) - (predict(mlmfit.log.steps, newdata = mlmtest)))^2)
plot(mlmfit.log.steps)
```

Lasso with log transformation (needs fixing)
```{r}
ridgefit <- glmnet(log(targdol)~slstyr + slslyr + sls2ago + slshist + 
    ordtyr + ordlyr + ord2ago + ord3ago, data = mlmtrain, alpha=1, lambda=seq(0,10,0.01))
summary(ridgefit)
```


```{r}
mlmfit.not <- lm(targdol ~ slsrecent:slshist + oldSales:slshist + ordrecent:ordhist + oldOrds:ordhist + recency, data=mlmtrain)
summary(mlmfit.not)
sum(((mlmtest$targdol) - (predict(mlmfit.not, newdata = mlmtest)))^2)
```

```{r}
hist(mlmtrain$targdol)
```

full model will have:
targdol~targdol+slstyr+slslyr+sls2ago+sls3ago+slshist+ordtyr+ordlyr+ord2ago+ord3ago+ordhist+oldOrds+oldSales+recency+slsrecent+ordrecent
```{r}
pfit <- glm(targdol~slsrecent+ordrecent+oldOrds+oldSales+recency, data = mlmtrain, family=poisson(log))
summary(pfit)
sum(((mlmtest$targdol) - (predict(pfit, newdata = mlmtest)))^2)
```

##Max sales dollars in the test model
```{r}
this <- newtest %>%
  arrange(desc(targdol)) %>%
  top_n(1000, targdol) %>%
  summarise(sum(targdol), mean(targdol))
```

```{r}
mlmtest %>%
  summarise_all(mean)

mlmtest %>%
  summarise_all(sd)
```

```{r}
hist(log(mlmtest$targdol))
```

```{r}
hist(log(mlmtrain$targdol))
```

```{r}
cor(mlmtrain)
```

